---
title: "CO2 effects on litter production"
author: "Beni Stocker"
date: "`r Sys.Date()`"
output:
  html_document:
    #theme: paper
    toc: true
    toc_float: true
    toc_depth: 2
# output:
#   pdf_document:
#     toc: true
#     toc_depth: 2
header-includes:
   - \usepackage{amsmath}
# bibliography: bibliography.bib
---

```{r, warning=FALSE}
library(readr)
library(dplyr)
library(metafor)   # see ?dat.curtis1998 for an example with CO2 manipulation data
library(ggplot2)
library(stringr)
```

# Data screening

For a first data screening want to do the following:

1. Get data from eCO2 experiments
2. Data availability per experiment
3. Data availability per variable
4. Long versus wide issue

Let's check first if the data is ready to use for these steps.

But before we start:

- Some variable names end with an underscore (`ANPP_`, `Asat_`, `SLA_`, `coarse_root_biomass_`, `LMA_`, `soil_NH4+_`, `N_uptake_`). Needed to delete that manually.
- Replaced lots of superscripted numbers (actually appeared as superscript when opening the CSV file in my text editor) with plain number (e.g., `m2`)
- Replaced units given like `g_m-2` with `g/m2`, or `kg_ha-1` with `kg/ha`, or `g/m2*yr` with `g/m2/yr` (assuming that's correct if that's the way it would be entered in R), `_g/m2` with `g/m2`, `cm2/m2*d` with `cm2/m2/d`, ... I'll leave it with that for now.
- Replaced `Vcmax_normalized_to_25°C` with `Vcmax25`, and `Jmax_normalized_to_25°C` with `Jmax25` (problematic degree sign)
- Removed weird symbol in units, e.g., `‘mol_CO2/m2/s` to `mol_CO2/m2/s`
- Removed line break in date string

**All that is saved as a new file `NewData_wide_CORRECTED2.csv`**

Load the corrected file, which is based on the one Kevin sent on 1.3.2019.
```{r, warning=FALSE}
library(readr)
library(dplyr)
df <- read_csv("~/data/gcme/data_received_190325/NewData_wide_CORRECTED2.csv") %>%
  
  # something is wrong in columns ambient_Sd, ambient_Se, elevated...
  mutate( ambient_Sd  = as.numeric(ambient_Sd),  ambient_Se  = as.numeric(ambient_Se), 
          elevated_Sd = as.numeric(elevated_Sd), elevated_Se = as.numeric(elevated_Se) )
  
  # XXX testing
  # filter(exp_nam == "SwissFACE_lolium2" & Data_type == "aboveground_biomass")
  # filter(exp_nam == "SwissFACE_trifolium2" & Data_type == "aboveground_biomass")

# save experiments names
df_experiments <- df %>% select(exp_nam, prev_name) %>% distinct()
```

This may be remediated automatically by using only columns that are distinct w.r.t. `valuevars` and `keyvars` (basically all columns in original data, except `ALIAS`).
```{r}
keyvars <- c("exp_nam", "factors", "treatment", "Data_type", "Unit", "Sampling_date", "Start_Year", "Year", "n_plots")
valuevars <-  c("ambient", "ambient_Se", "ambient_Sd", "elevated", "elevated_Se", "elevated_Sd")

df <- df %>% 
  distinct_at(vars(one_of(keyvars, valuevars))) %>%
  mutate( id=1:n() )   # create new ID key (before this was 'ALIAS')
```


# Data selection

## General filters

We filter the data based on the following criteria:

- Only data from experiments is used where CO2 was the only manipulated factor (data from `treatment = "c"`)
- Only data from experiments where data from at least two years is available, based on column `Year` in the data table.
- Only data from Open Top Chambers or Free Air CO2 Enrichment CO2 experiments: fumigation type is either `"OTC"` or `"FACE"`. 

```{r warning=FALSE}
## Determine all experiments that have more than 1 year data
list_exp_gt1yr <- df %>% 
  filter(!is.na(Year)) %>% 
  group_by(exp_nam) %>% 
  summarise(nyears=max(Year)) %>% 
  filter(nyears>1) %>% 
  select(exp_nam) %>% 
  unlist() %>% 
  unname()

df_c <- df %>%
  
  # ## Take this info from experiments below
  # select(-Fumigation_type, -Vegetation_type) %>% 
  
  ## Add prev_name back
  left_join( df_experiments, by = "exp_nam") %>% 
  
  ## filter experiments with only manipulated CO2
  ## (no other factors manipulated, strong reduction of data)
  filter(treatment=="c") %>% 
  
  ## More than 1 year data
  filter(exp_nam %in% list_exp_gt1yr) %>% 
  
  ## Combine with experiments meta info
  left_join( 
    read_csv("~/data/gcme/data_received_190325/table_var_exp_names_experiments.csv") %>% 
      select(prev_name, Fumigation_type=my_fumigation_type, Vegetation_type),
    by = c("prev_name")
  ) %>% 
  
  ## Filter only Fumigation_type OTC or FACE
  filter( Fumigation_type %in% c("OTC", "FACE") ) %>%
  
  {.}
```


<!-- Add `my_varnam` to the data table, defined by `my_varnam = ifelse( varnam %in% c("anpp", "anpp_bm"), "my_anpp", my_varnam )`. WARNING: This assumes that variable names 'Data_type' are identical in old and new dataset (manually replaced white spaces with underscores in table_var_exp_names_anpp.csv, first) -->
<!-- ```{r, warning=FALSE} -->
<!-- df_c <- df_c %>%  -->
<!--   mutate( varnam=NA ) %>%  -->
<!--   left_join( select( df_anpp, -prev_name ), by=c("exp_nam", "Data_type") ) %>%  -->
<!--   distinct(id, .keep_all=TRUE) %>%  -->
<!--   rename( varnam = varnam.x ) %>%  -->
<!--   rowwise() %>%  -->
<!--   mutate( varnam = ifelse(!is.na(varnam.y), varnam.y, varnam ) ) %>%  -->
<!--   select( -varnam.y ) %>%  -->
<!--   mutate( my_varnam = NA ) %>%  -->
<!--   mutate( my_varnam = ifelse( varnam %in% c("anpp", "anpp_bm"), "my_anpp", my_varnam ) ) -->
<!-- ``` -->



## Litter

Let's pool various types of data to analyse response ratios of litter production and mass, interpreted here as a proxy for total SOC input.

Litter: `my_varnam = my_litter`, and `varnam %in% c("litter_prod", "litter_mass", "litter_prod_rel")`.

Collect information about experiments providing data for any of `varnam %in% c("litter_prod", "litter_mass", "litter_prod_rel")` from file `"table_var_exp_names_litter.csv"`. See procedure [here](https://docs.google.com/document/d/11ghws59q6Ts6j6GK6yNqQBvilrNojRN_aC7c41-LrI0/edit?usp=sharing). Add additional information from the experiments table (`table_var_exp_names_experiments.csv`) and from the data frame `df_experiments`, which is used to associate new and old experiments names.
```{r, warning=FALSE}
df_litter <- read_csv("~/data/gcme/data_received_190325/table_var_exp_names_litter.csv") %>% 
  select(prev_name=expnam, varnam, Data_type, Source_Reference) %>%
  distinct() %>% 
  filter(varnam %in% c("litter_prod", "litter_mass", "litter_prod_rel") ) %>% 
  
  ## add 'exp_nam' (new experiment name) from data table
  left_join( df_experiments, by="prev_name" ) %>% 
  
  ## filter only experiments selected by general filters above
  filter( exp_nam %in% unique(df_c$exp_nam) ) %>% 
  
  ## add experiments meta information from experiments table
  left_join( 
    read_csv("~/data/gcme/data_received_190325/table_var_exp_names_experiments.csv") %>% 
      select(prev_name, Fumigation_type=my_fumigation_type, Cquantity_Control, Cquantity_Treatment, Vegetation_type, Plants_specs, Start_Year, End_Year, Remarks),
    by = c("prev_name")
    ) %>% 
  
  ## String replace in CSV file
  mutate( Data_type = str_replace_all(Data_type, " ", "_") )
```

Add `my_varnam` to the data table, defined by `my_varnam = ifelse( varnam %in% c("litter_prod", "litter_mass", "litter_prod_rel"))`.
```{r, warning=FALSE}
df_c <- df_c %>%
  mutate( varnam=NA ) %>% 
  mutate( my_varnam = NA ) %>%
  left_join( select( df_litter, exp_nam, Data_type, varnam_unified=varnam ), by=c("exp_nam", "Data_type") ) %>%
  mutate( varnam = ifelse( is.na(varnam), varnam_unified, varnam ) ) %>%
  mutate( my_varnam = ifelse( varnam %in% c("litter_prod", "litter_mass", "litter_prod_rel"), "my_litter", my_varnam ) ) %>%
  select(-varnam_unified)
```


## For labeling

Prepare variable name association for labelling plots
```{r}
df_varnams <- tibble(
  my_varnam = c("my_litter"),
  my_lab    = c("Litter")
)
```


# Data analysis 

Test analysis and plot of ANPP data. Calculate the response ratio of ANPP (mean and variance) for each experiment. To get that, we first need to calcuate the means and standard deviation for the ambient and elevated levels, pooling multiple measurements (years, sampling dates), each given with mean $\mu_i$, number $N_i$ (replicates/plots), and standard deviation $\sigma_i$ or standard error. For the function `metafor::escalc()`, we need standard deviations ($SD$). Calculate them for those rows where only standard errors $SE$ are given as:
$$
SD = SE \sqrt{N}
$$

Now, do the meta-analysis and plot results.
```{r, warning=FALSE}
library(metafor)   # see ?dat.curtis1998 for an example with CO2 manipulation data

## aggregate by variable and experiment, pooling multiple years, sampling dates, and plots/replicates and calculate log response ratio
df_c_sub <- df_c %>%         
  
  ## Here only for my variables, selected as described above 
  ## (leaving out "my_anet"; omitting "my_nleach" and "my_n2o" because only one experiment available)
  filter(my_varnam %in% c("my_litter")) %>%
  
  # get standard deviation for all data
  mutate( my_ambient_sd = ambient_Sd, my_elevated_sd = elevated_Sd ) %>%
  rowwise() %>% 
  mutate( my_ambient_sd   = ifelse( is.na(my_ambient_sd),  ambient_Se  * sqrt(n_plots), my_ambient_sd ),
          my_elevated_sd  = ifelse( is.na(my_elevated_sd), elevated_Se * sqrt(n_plots), my_elevated_sd )) %>%

  ## Get logarithm of response ratio and its variance
  metafor::escalc( 
    measure = "ROM", 
    m1i = elevated, sd1i = my_elevated_sd, n1i = n_plots, 
    m2i = ambient,  sd2i = my_ambient_sd,  n2i = n_plots, 
    data=., 
    append = TRUE, var.names = c("logr", "logr_var") ) %>% 
  as_tibble() %>% 
  mutate( logr_se = sqrt(logr_var)/sqrt(n_plots) )

## pool all measurements (multiple years, sampling dates and plots) by variable and experiment for meta analysis
df_c_agg <- df_c_sub %>% 
  filter(!is.na(logr_var) & !is.na(logr)) %>% 
  select(-id) %>% # this is necessary because the dumb function agg() uses my 'id' otherwise
  mutate( id = paste(exp_nam, my_varnam, sep="_XXX_")) %>% 
  MAd::agg( id = id, es = logr, var = logr_var, n.1 = n_plots, n.2 = n_plots, cor = 1.0, method = "BHHR", data = . ) %>% 
  as_tibble() %>% 
  mutate( id = str_split(id, "_XXX_") ) %>% 
  mutate( exp_nam = purrr::map_chr(id, 1),
          my_varnam = purrr::map_chr(id, 2) ) %>% 
  select(exp_nam, my_varnam, es, var) %>% 

  ## add number of plots column and my_varnam
  left_join( df_c_sub %>% 
               group_by( exp_nam, my_varnam ) %>%
               summarise( n_plots = sum(n_plots) ) %>% 
               select( exp_nam, my_varnam, n_plots ),
             by = c("exp_nam", "my_varnam") ) %>% 
  rename( logr = es, logr_var = var ) %>% 
  mutate( logr_se = sqrt(logr_var)/sqrt(n_plots) ) %>% 
  left_join( df_varnams, by = "my_varnam" ) %>% 
  
  ## filter NA for exp_nam due to unidentified experiment name in soil decomposition dataset
  filter(exp_nam!="NA" & !is.na(exp_nam))

## aggregate by variable, doing a meta-analysis of the log response ratios, with experiment as random factor (method="DL")
agg_meta <- function(df, groupvar){
  
  out_meta <- df %>% dplyr::filter(my_varnam==eval(parse_character(groupvar))) %>% 
    
    # main meta analysis function call, adjusted step size (see http://www.metafor-project.org/doku.php/tips:convergence_problems_rma)
    # metafor::rma( logr, logr_var, method = "REML", slab = exp_nam, control = list(stepadj=0.3), data = . )
    metafor::rma.mv( logr, logr_var, method = "REML", random = ~ 1 | exp_nam, slab = exp_nam, control = list(stepadj=0.3), data = . )
  
  # transform back
  out_meta_scaled <- predict( out_meta, transf=exp )
  
  df_box <- tibble(
    my_varnam=groupvar, 
    middle = out_meta$b[1,1], 
    ymin   = out_meta$ci.lb, 
    ymax   = out_meta$ci.ub,
    
    middle_scaled = out_meta_scaled$pred, 
    ymin_scaled   = out_meta_scaled$ci.lb, 
    ymax_scaled   = out_meta_scaled$ci.ub
  )
  return(list(df_box=df_box, out_meta=out_meta))
}

# do meta-analysis on all variables
varlist <- unique(df_c_agg$my_varnam)
list_meta  <- purrr::map(as.list(varlist), ~agg_meta(df_c_agg, .))
df_metabox <- purrr::map_dfr(list_meta, "df_box") %>% left_join( df_varnams, by = "my_varnam" )
names(list_meta) <- varlist
```

Plot dots and my box.
```{r}
library(ggplot2)
df_c_agg %>%
  # arrange(logr) %>% 
  mutate( my_lab = factor(my_lab, levels=rev(c("Litter")))) %>% 
  ggplot( aes(x=my_lab, y=logr)) +
  geom_jitter( color = rgb(0,0,0,0.3), aes( size = 1/logr_se ), position = position_jitter(w = 0.2, h = 0) ) +
  geom_crossbar( data = df_metabox, aes(x=my_lab, y=middle, ymin=ymin, ymax=ymax), fill = "grey80", alpha = 0.6, width = 0.5 ) +
  geom_hline( yintercept=0.0, size=0.5 ) +
  labs(x="", y="Log Response Ratio", size=expression(paste("Error"^{-1}))) +
  coord_flip() +
  ylim(-1,1) 

ggsave("fig/meta_litter_myboxes.pdf", width=6, height=4)
```

Number of data points (plot-level measurements) per variable:
```{r}
df_c_agg %>% 
  group_by(my_lab) %>% 
  summarise(n_plots=sum(n_plots), n_exp=n()) %>% 
  rename_("Variable"="my_lab", "N plots"="n_plots", "N experiments"="n_exp") %>% 
  knitr::kable()
```

## Forest plots

And the forest plots:
```{r}
forest_byvar <- function(list_meta, varnam_lab){
  par(mar=c(4,4,1,2))
  out_forest <- metafor::forest(list_meta$out_meta, xlab="Log Response Ratio", mlab="", xlim=c(-1,1), cex=0.5)
  text(out_forest$xlim[1], out_forest$ylim[2], varnam_lab,  pos=4, cex=0.7, font=2 )
}

df_varnams_sub <- df_varnams %>% filter(!(my_varnam %in% c("my_anet", "my_n2o", "my_nleach")))
purrr::map(
  as.list(1:nrow(df_varnams_sub)), 
  ~forest_byvar(list_meta[[df_varnams_sub$my_varnam[.]]], df_varnams_sub$my_lab[.])
  )
```

This shows the log-response ratio of ANPP, each dot for one experiment, multiple years, sampling dates, and plots pooled. The size of the dots represents the standard error of the response ratio (the smaller, the bigger). The box represents the meta-analysis mean and confidence interval of the log response ratio.

# Save data

Write subsetted data to files for Cesar.
```{r}
df_c_sub %>% 
  write_csv(path = "data/data_gcme_litter.csv")

df_c_agg %>% 
   write_csv(path = "data/data_agg_gcme_litter.csv")
```

