---
title: "Simplest CN-model"
author: "Beni Stocker"
date: "`r Sys.Date()`"
output:
  html_document:
    # theme: cosmo  #paper
    toc: true
    toc_float: true
    toc_depth: 2
# output:
#   pdf_document:
#     toc: true
#     toc_depth: 2
header-includes:
   - \usepackage{amsmath}
# bibliography: bibliography.bib
---

\newcommand{\rcton}{r_{\text{C:N}}}
\newcommand{\nacq}{N_{\text{acq}}}
\newcommand{\netmin}{N_{\text{min}}}
\newcommand{\nin}{N_{\text{in}}}


One may assume that the general dynamics of the coupled C and N cycling in terrestrial ecosystems can be described by a small list of processes (functions). The simplest CN-model ("sCN-model") embodies the following relationships:

- The plant (divided into an above and a belowground compartment) balances the acquisition of C and N in order to satisfy its stoichiometric requirements (fixed: $r_{C:N}$), imposed by new growth (*balanced growth condition*). It tends to achieve a balance between supply of N through acquisition pathways and the demand imposed by new growth and its respective C:N ratio.
$$
\nacq (C_r) = y/\rcton [ P(C_l) - r(C_l+C_r) ] \; \; \; \; \; \; (1)
$$
$P$ is the C assimilation rate and is a function of leaf mass ($C_l$), $y$ is the growth efficiency, $r$ a respiration coefficient, implying that plant respiration scales with its size, $C_l$ and $C_r$ are the leaf carbon and root pool sizes respectively, and $N_{\text{acq}}$ is the the N acquisition flux and is a function of the root mass ($C_r$). 

- The size of the above-ground pool $C_l$ (each pool consists of C and N) determines the assimilation rate of C ($P$) and has declining marginal returns towards increasing $C_l$.
$$
P(C_l) = I \varepsilon \left(1 - e^{-k_b \sigma C_{\text{l}}} \right)
$$
Here, $I$ is PPFD, $\varepsilon$ is the light use efficiency, $k_b$ is the light extinction parameter, and $\sigma$ is the specific leaf area.
To simplify the mathematics, I've tried to model this with Michaelis-Menten as an alternative:
$$
P(C_l) = I \varepsilon \frac{C_l}{C_l + K_P}
$$

```{r echo=FALSE}
prod <- function( cleaf, ppfd, lue ){
  # kbeer         = 0.5
  # sla           = 0.1
  #prod <- ppfd * lue * ( 1.0 - exp( - kbeer * sla * cleaf ) )
  
  kmm_prod <- 100
  prod <- ppfd * lue * cleaf / ( cleaf + kmm_prod )
  return(prod)
}
## Example
curve( prod( x, ppfd = 100, lue = 1 ), from = 0, to = 350, xlab = expression(italic(C[l])), ylab = expression(italic(P)) )
```


- The size of the below-ground pool ($C_r$) determines the fraction of net N mineralisation $\netmin$ that is acquired by the plant ($\nacq$). As for $P$, the function $f(C_r)$ represents declining marginal returns in $\nacq$ with increasing $C_r$:
$$
\nacq = f(C_r) \netmin   \\
f(C_r) = (1 - u)  (1 - e^{-k_r  C_r } ) 
$$
Here, $u$ is the unaccessible fraction of net mineralisation, i.e. N losses that are not subject to plant control.
Again, in order to simplify mathematics, I've tried to model this with Michaelis-Menten as an alternative:
$$
f(C_r) = (1-u) \frac{C_r}{C_r + K_N}
$$

```{r echo=FALSE}
f_noloss <- function( croot ){
  f_unavoidable = 0.1
  k_noloss      = 100
  out <- (1.0 - f_unavoidable) * croot / (croot + k_noloss )
  # f_avl <- min( 1.0, croot / 30 )
  # out   <- (1.0 - f_unavoidable) * (1.0 - exp( - 0.01 * croot ) )
  return(out)
}
## Example
netmin <- 1.0
curve( netmin * f_noloss(x), from = 0, to = 350, ylim=c(0,1), xlab = expression(italic(C[r])), ylab = expression(italic(N[acq])) )
abline( h=(1-0.1), lty=3 )
abline( h=netmin, lty=2 )
```

## Example

For given $I$, $\varepsilon$, $N_{\text{min}}$, a balance of above-to-belowground pool size $a = C_l:C_r$ can be found that satisfies Eq. (1). The imbalance can be expressed as
$$
f((1-a)C)\netmin = y/\rcton\; [ P(aC) - rC ]
$$
Here, $C=C_l+C_r$ and $a=C_l/C$. Below is an example plot for the imbalance term as a function of $a$. 

```{r echo=FALSE}
eval_cnbalance <- function( f_ag, ctot, netmin, lue, ppfd ){
  eff  = 0.7
  resp = 0.1
  r_cton_plant = 30

  ## version where not all aboveground C is leaves
  # out <- (( eff  * ( prod( f_leaf_of_c_ag_tot( f_ag * ctot ) * f_ag * ctot, ppfd, lue ) - resp * ctot ) ) /
  #        ( netmin * f_noloss( (1.0 - f_ag) * ctot ) ) ) - r_cton_plant

  ## version where all aboveground C is leaves
  # out <- (( eff  * ( prod( f_ag * ctot, ppfd, lue ) - resp * ctot ) ) /
  #        ( netmin * f_noloss( (1.0 - f_ag) * ctot ) ) ) - r_cton_plant

  ## this works much better than the ratio calculated above
  supply <- f_noloss((1.0 - f_ag) * ctot ) * netmin
  demand <- eff * ( prod( f_ag * ctot, ppfd, lue ) - resp * ctot ) / r_cton_plant
  out <- supply - demand

  return( out )
}

findroot_eval_cnbalance <- function( ... ){

  if ( eval_cnbalance( 0.0, ... ) < 0.0 ){
    ## all to roots
    fleaf <- 0.0
  } else if ( eval_cnbalance( 1.0, ... ) > 0.0 ){
    ## all to leaves
    fleaf <- 1.0
  } else {
    ## find allocation fraction to leaves so that return next year matches (eff * r_cton_plant)
    out_root <- uniroot( function(x) eval_cnbalance( x, ... ), interval=c(0,1.0) )
    fleaf <- out_root$root
  }
  return(fleaf)
}

lue  = 1
# ppfd = 60 # for example with exponentials
ppfd = 100 # for example with MM
netmin = 0.8
ctot = 350

## Imbalance as a function of the C_l:C
curve( eval_cnbalance( x, ctot, netmin, lue, ppfd ), from = 0, to = 1.0, col="red", xlab=expression(italic(C[l]/C)), ylab=expression(Imbalance) ) #, ylim=c(-20,100)
abline( h=0, lty=3 )
out_root <- findroot_eval_cnbalance( ctot, netmin, lue, ppfd )  # uniroot( function(x) eval_cnbalance( x, ctot, netmin, lue, ppfd ), interval=c(0,1.0) )
abline( v = out_root, lty = 2 )
```

## Steady-state solution

In steady state, the total stock of N in the soil is constant. This implies that net mineralisation equals total N inputs through litterfall and N deposition (assuming deposited N bypasses plants). For the sake of simplicity, N fixation and resorption are ignored here.
$$
\netmin = \frac{C}{\rcton\tau} + \nin
$$
Here, $\tau$ is the residence time of C and N in the plant pool; $\rcton$ is the C:N ratio of plant biomass; and $\nin$ is the external input of N, subsuming atmospheric deposition and weathering. Using this, the steady-state solution of the coupled soil-plant system can be found, without relying on prescribed net mineralisation rates, with the balanced-growth condition from above:
$$
 f((1-a)C)(\frac{C}{\rcton\tau} +\nin) =  y / \rcton \; [ P(aC) - rC ]\; \; \; \;\; \; \; \;(2)
$$
and the plant C blance (C assimilation equals respiration plus new growth, replacing litterfall in steady-state):
$$
P(aC) - (\tau/y + r)\;C = 0 \; \; \; \; \; \; \; \;(3)
$$
This set of two equations should in theory allow us to cancel $C$ and solve for $a$ as a function of environmental conditions: $a = f(\varepsilon, I, \nin)$. Note that $\varepsilon$ is not treated here as a parameter because it's itself a function of environmental conditions. Let's not try to get the analytical solution here. It will involve Weibull functions anyways, and these have to be solved numerically. Let's just throw it at `nleqslv()`.

```{r}
library(nleqslv)

## Equation system as a function returning a vector of two values, both of which are to be set to zero by the optimiser
fn <- function( par ){
  eff  <- 0.7
  resp <- 0.1
  r_cton_plant <- 30
  tau_plant <- 10
  n_in <- 1.0

  out <- numeric(2)
  ctot <- par[1]
  f_ag <- par[2]
  
  out[1] <- eval_cnbalance( f_ag, ctot, ( ctot/(r_cton_plant * tau_plant) + n_in), lue, ppfd )
  out[2] <- prod( f_ag * ctot, ppfd, lue ) - ctot * ( tau_plant/eff + resp )
  
  return(out)
}

steadystate <- nleqslv( c(500, 0.5), fn, control = list( allowSingular=TRUE ) )

## test
ctot <- steadystate$x[[1]]
f_ag <- steadystate$x[[2]]
  eff  <- 0.7
  resp <- 0.1
  r_cton_plant <- 30
  tau_plant <- 10
  n_in <- 1.0
  print(paste("should be zero:", eval_cnbalance( f_ag, ctot, ( ctot/(r_cton_plant * tau_plant) + n_in), lue, ppfd ) ))
  print(paste("should be zero:", prod( f_ag * ctot, ppfd, lue ) - ctot * ( tau_plant/eff + resp ) ))
```

### My problem

Something is apparently not right here. I've tried to solve this system of two non-linear equations numerically but the solution I get is nonsense. A problem might be that I cannot constrain the solutions to a (phyically meaningful) interval, $0<a<1$ in this case. But the nature of the problem shouldn't require this anyways. So I am asking myself if the second equation above (Eq. 3) is appropriate or if something else is wrong.

So far, I haven't attempted to analytically solve this. Feel free if you like...

```{r eval=FALSE, echo=FALSE}

# ## enther this in wolfram alpha:
# ((1 - u) * (1 - y) * x / ((1 - y) * x + l)) * ( x/(s*t) + d) - e * ( (I * p * x * y / (x * y + k)) - r * x ) / s = 0
# (I * p * x * y / (x * y + k)) - (t/e + r) * x = 0

lue  <<- 1
ppfd <<- 100 # for example with MM
eff  <<- 0.7
resp <<- 0.1
r_cton_plant <<- 30
f_unavoidable <<- 0.1
k_noloss <<- 100
kmm_prod <<- 100
n_in <<- 1.0
tau_plant <<- 10

fn <- function( x ){
  
  ctot <- x[1]
  f_ag <- x[2]
  out <- numeric(2)
  
  supply <- f_noloss((1.0 - f_ag) * ctot ) * netmin
  demand <- eff * ( prod( f_ag * ctot, ppfd, lue ) - resp * ctot ) / r_cton_plant
  out[1] <- supply - demand

  out[2] <- prod( f_ag * ctot, ppfd, lue ) - (tau_plant / eff + resp) * ctot

  return(out)
}

## Does not work, maybe I'm looking in an implausible range of values
library(pracma)
solution_pracma <- pracma::fsolve( fn, c(230.8476, 0.8366655) )

library(nleqslv)
solution_nleqslv <- nleqslv::nleqslv( c(230.8476, 0.8366655), fn )

print( "Steady-state solution:" )
print( paste0( "C_l: ", solution_pracma$x[1] * solution_pracma$x[2] ))
print( paste0( "C_r: ", solution_pracma$x[1] * (1-solution_pracma$x[2]) ))
```

<!-- ```{r} -->
<!-- c_wood <- function( ctot ){ -->
<!--   par = 50 -->
<!--   sqrt( ctot^2 + par ) - sqrt(par) -->
<!-- } -->

<!-- c_leaf <- function( ctot ){ -->
<!--   kl = 0.3 -->
<!--   c_leaf_max = 20 -->
<!--   c_leaf_max * (1 - exp(-kl*ctot)) -->
<!-- } -->

<!-- f_leaf_of_c_ag_tot <- function( c_ag_tot ){ -->
<!--   ifelse( c_ag_tot==0, 1.0, c_leaf( c_ag_tot ) / (c_leaf( c_ag_tot ) + c_wood( c_ag_tot )) ) -->
<!-- } -->
<!-- ``` -->

<!-- ## Static CN-model for one state------------------------------- -->
<!-- lue  = 1 -->
<!-- ppfd = 100 -->
<!-- navl = 0.8 -->
<!-- ctot = 350 -->

<!-- curve( c_wood(x), from = 0, to = 50 ) -->
<!-- curve( c_leaf(x), from = 0, to = 50, add = TRUE ) -->
<!-- curve( f_leaf_of_c_ag_tot(x), from = 0, to = ctot ) -->

<!-- curve( prod( f_leaf_of_c_ag_tot( x ) * x, ppfd, lue ), from = 0, to = ctot ) -->
<!-- curve( f_noloss, from = 0, to = ctot*2 ) -->

<!-- ## Visualise imbalance curve -->
<!-- par(mfrow=c(1,1)) -->
<!-- curve( eval_cnbalance( x, ctot, navl, lue, ppfd ), from = 0, to = 1.0, col="red", ylim=c(-20,100) ) -->
<!-- abline( h=0, lty=3 ) -->
<!-- out_root <- uniroot( function(x) eval_cnbalance( x, ctot, navl, lue, ppfd ), interval=c(0,1.0) ) -->
<!-- abline( v = out_root$root, lty = 2 ) -->

<!-- ## production curve -->
<!-- curve( prod( f_leaf_of_c_ag_tot( x ) * x, ppfd, lue ), from = 0, to = ctot ) # x is total aboveground -->
<!-- points( out_root$root * ctot, prod( f_leaf_of_c_ag_tot( out_root$root * ctot ) * out_root$root * ctot, ppfd, lue ), pch=16, col="red" ) -->

<!-- ## N uptake curve -->
<!-- curve( navl * f_noloss( x ), from = 0, to = ctot ) # x is total belowground -->
<!-- points( (1.0 - out_root$root) * ctot, navl * f_noloss( (1.0 - out_root$root) * ctot ), pch=16, col="red" ) -->


## Dynamic CN-model

Above equations can also be implemented in a simple dynamical model that additionally accounts for soil C and N turnover. The results from a spinup to equilibrium look like this (suggesting that there is a physically meaningful solution). In this case (last plot at the very bottom), the ratio of $C_l/C$ is around 0.83. 

```{r echo=FALSE}
lue  <- 1
ppfd <- 70 # for example with MM
eff  <- 0.7
resp <- 0.1
r_cton_plant <- 30
f_unavoidable <- 0.1
k_noloss <- 100
kmm_prod <- 100
n_in <- 1.0
tau_plant <- 10

## other parameters defined above
r_cton_soil <- 10
tau_soil_c <- 50
tau_soil_n <- 150
tau_labl   <- 0.5

r_ntoc_plant <- 1/r_cton_plant
r_ntoc_soil  <- 1/r_cton_soil

## Simulation settings
ntsteps <- 2000

## initialise output variables
out_cplant_ag <- rep(NA, ntsteps)
out_nplant_ag <- rep(NA, ntsteps)
out_cplant_bg <- rep(NA, ntsteps)
out_nplant_bg <- rep(NA, ntsteps)
out_csoil     <- rep(NA, ntsteps)
out_nsoil     <- rep(NA, ntsteps)
out_clabl     <- rep(NA, ntsteps)
out_nlabl     <- rep(NA, ntsteps)

## plant, aboveground
cplant_ag <- 150
nplant_ag <- cplant_ag * r_ntoc_plant

## plant, belowground
cplant_bg <- 150
nplant_bg <- cplant_bg * r_ntoc_plant

## soil
csoil <- 1000
nsoil <- csoil * r_ntoc_plant * tau_soil_n/tau_soil_c

## plant labile pools
nlabl <- 0
clabl <- 0

## to avoid numerical oscillation
calc_turnover <- function( c0, tau, dt = 1.0 ){
  turnover <- c0 * (1.0 - exp(-1/tau * dt ) )
  return(turnover)
}

for (it in 1:ntsteps){

  # ## step increase in light use efficiency ~ CO2 fertilisation
  # if (it==1000) lue <- 1.2 * lue
  
  ## Soil turnover
  csoil_turnover <- calc_turnover( csoil, tau_soil_c )
  csoil          <- csoil - csoil_turnover

  nsoil_turnover <- calc_turnover( nsoil, tau_soil_n )
  nsoil          <- nsoil - nsoil_turnover

  ## Net mineralisation
  netmin         <- nsoil_turnover + n_in

  ## actual C and N acquisition with current plant
  # clabl <- prod( f_leaf_of_c_ag_tot( cplant_ag ) * cplant_ag, ppfd, lue ) #+ clabl
  clabl <- prod( cplant_ag, ppfd, lue ) #+ clabl
  nlabl <- f_noloss( cplant_bg ) * netmin #+ nlabl

  # ## short-cut: by-passing soil
  # nlabl <- f_noloss( cplant_bg ) * r_ntoc_plant * (cplant_bg + cplant_ag) / tau_plant  #+ nlabl

  ## limit allocatable C depending on labile N, c_alloc is used also for growth respiration
  c_alloc <- min( nlabl / (eff * r_ntoc_plant), clabl )

  # ##----------
  # netmin <- 0.8 # xxx test
  # ##----------

  ## Plant turnover, needs to be after acquisition and before new balance evaluation
  cturnover_ag <- calc_turnover( cplant_ag, tau_plant )
  cplant_ag    <- cplant_ag - cturnover_ag

  nturnover_ag <- calc_turnover( nplant_ag, tau_plant)
  nplant_ag    <- nplant_ag - nturnover_ag

  cturnover_bg <- calc_turnover( cplant_bg, tau_plant )
  cplant_bg    <- cplant_bg - cturnover_bg

  nturnover_bg <- calc_turnover( nplant_ag, tau_plant)
  nplant_ag    <- nplant_ag - nturnover_bg

  csoil <- csoil + cturnover_ag + cturnover_bg
  nsoil <- nsoil + nturnover_ag + nturnover_bg

  ## Get optimal fraction to leaves
  fleaf <- findroot_eval_cnbalance( (c_alloc * eff + cplant_ag + cplant_bg), netmin, lue, ppfd )
  # print(fleaf)

  # ##----------
  # ## Visualise imbalance curve
  # ctot <- (cplant_ag + cplant_bg + eff * c_alloc)
  # par(mfrow=c(1,1))
  # curve( eval_cnbalance( x, ctot, netmin, lue, ppfd ), from = 0, to = 1.0, col="red", ylim=c(-20,100) )
  # abline( h=0, lty=3 )
  # abline( v = fleaf, lty = 2 )

  # ## production curve
  # curve( prod( f_leaf_of_c_ag_tot( x ) * x, ppfd, lue ), from = 0, to = ctot ) # x is total aboveground
  # points( fleaf * ctot, prod( f_leaf_of_c_ag_tot( fleaf * ctot ) * fleaf * ctot, ppfd, lue ), pch=16, col="red" )

  # ## N uptake curve
  # curve( netmin * f_noloss( x ), from = 0, to = ctot ) # x is total belowground
  # points( (1.0 - fleaf) * ctot, netmin * f_noloss( (1.0 - fleaf) * ctot ), pch=16, col="red" )
  # ##----------

  ## Additional check only necessary for legacy allocation setup (Version A)
  # if ( eval_cnbalance( 0, ctot, netmin, lue, ppfd ) > 0.0 ){
  #   ## all to roots
  #   fleaf <- 0.0
  # } else if ( eval_cnbalance( 1, ctot, netmin, lue, ppfd ) < 0.0 ){
  #   ## all to leaves
  #   fleaf <- 1.0
  # } else {
  #   ## find allocation fraction to leaves so that return next year matches (eff * r_cton_plant)
  #   out_root <- uniroot( function(x) eval_cnbalance( x, (cplant_ag + cplant_bg + eff * c_alloc), netmin, lue, ppfd ), interval=c(1e-12,1.0) )
  #   fleaf <- out_root$root
  # }

  ## update labile pools
  clabl <- clabl - c_alloc
  nlabl <- nlabl - eff * c_alloc * r_ntoc_plant

  ## labile pool decay
  clabl <- clabl - calc_turnover( clabl, tau_labl)
  nlabl <- nlabl - calc_turnover( nlabl, tau_labl)

  ##---------------------------------------------------------------------
  # ## VERSION A: fleaf only acts on new growth
  # ## allocate C, N follows from r_ntoc_plant
  # cturnover_ag <- cplant_ag / tau_plant
  # c_alloc_ag   <- fleaf * c_alloc * eff
  # cplant_ag    <- cplant_ag + c_alloc_ag - cturnover_ag

  # nturnover_ag <- nplant_ag / tau_plant
  # n_alloc_ag   <- c_alloc_ag * r_ntoc_plant
  # nplant_ag    <- nplant_ag + n_alloc_ag - nturnover_ag

  # cturnover_bg <- cplant_bg / tau_plant
  # c_alloc_bg   <- (1.0 - fleaf) * c_alloc * eff
  # cplant_bg    <- cplant_bg + c_alloc_bg - cturnover_bg

  # nturnover_bg <- nplant_bg / tau_plant
  # n_alloc_bg   <- c_alloc_bg * r_ntoc_plant
  # nplant_bg    <- nplant_bg + n_alloc_bg - nturnover_bg
  ##---------------------------------------------------------------------

  ##---------------------------------------------------------------------
  ## VERSION B: fleaf redistributes entire plant (to avoid legacy in mal-allocation)
  ## allocate C, N follows from r_ntoc_plant
  cplant_ag    <- fleaf * (c_alloc * eff + cplant_ag + cplant_bg)
  cplant_bg    <- (1.0 - fleaf) * (c_alloc * eff + cplant_ag + cplant_bg)
  ##---------------------------------------------------------------------


  ## gater output variables
  out_cplant_ag[it] <- cplant_ag
  out_nplant_ag[it] <- nplant_ag
  out_cplant_bg[it] <- cplant_bg
  out_nplant_bg[it] <- nplant_bg
  out_csoil[it]     <- csoil
  out_nsoil[it]     <- nsoil
  out_clabl[it]     <- clabl
  out_nlabl[it]     <- nlabl

}
##---------------------------------------------------------
plot( 1:ntsteps, out_cplant_ag, type = "l" )
plot( 1:ntsteps, out_cplant_bg, type = "l" )
plot( 1:ntsteps, out_csoil, type = "l" )
plot( 1:ntsteps, out_nsoil, type = "l" )
plot( 1:ntsteps, out_clabl, type = "l" )
plot( 1:ntsteps, out_nlabl, type = "l" )
plot( 1:ntsteps, out_cplant_ag/(out_cplant_bg+out_cplant_ag), type = "l" )

print("Steady state f_ag (fraction of aboveground to total plant C):")
print( cplant_ag[ntsteps]/(cplant_bg[ntsteps]+cplant_ag[ntsteps]) )

print("Steady state total plant C:")
print( cplant_bg[ntsteps]+cplant_ag[ntsteps] )

print("Steady state net mineralisation:")
print(netmin)

print("Expected steady state net mineralisation:")
print( (cplant_ag + cplant_bg) / (r_cton_plant * tau_plant) + n_in )
```
