---
title: "Simplest CN-model"
author: "Beni Stocker"
date: "`r Sys.Date()`"
output:
  html_document:
    # theme: cosmo  #paper
    toc: true
    toc_float: true
    toc_depth: 2
# output:
#   pdf_document:
#     toc: true
#     toc_depth: 2
header-includes:
   - \usepackage{amsmath}
# bibliography: bibliography.bib
---

\newcommand{\rcton}{r_{\text{C:N}}}
\newcommand{\nacq}{N_{\text{acq}}}
\newcommand{\netmin}{N_{\text{min}}}
\newcommand{\nin}{N_{\text{in}}}

## Theory

One may assume that the general dynamics of the coupled C and N cycling in terrestrial ecosystems can be described by a small list of processes (functions). The simplest CN-model ("sCN-model") embodies the following relationships:

### Balance
The plant (divided into an above and a belowground compartment) balances the acquisition of C and N in order to satisfy its stoichiometric requirements (fixed N:C ratio in all tissues: $r$, no resorption), imposed by new growth (*balanced growth condition*). When a balance is achieved, the supply of N is equal to its demand:
$$
N_\text{supply} = N_\text{demand}
$$

### N supply

The N supply $N_\text{acq}$ can be described as a concave function of the total belowground C allocation $C_\text{bg}$. $N_\text{acq}$ reaches an asymptote, given by the amount of N that becomes available from net mineralisation and other external N inputs (e.g., atmospheric deposition), termed $N_0$. The common functional form of this relationship is a Monod equation:
$$
N_\text{supply} = N_0 \frac{C_\text{bg}}{C_\text{bg} + K_R}
$$
$K_R$ is the root-half saturation constant.

```{r echo=FALSE, message=FALSE, warning=FALSE}
library(dplyr)
library(ggplot2)

f_supply <- function( cbg, n0, f_unavoidable=0.0 ){
  kr <- 100
  (1.0 - f_unavoidable) * n0 * cbg / (cbg + kr )
}
df_nsupply <- tibble( cbg = seq(0, 350, length.out = 100) ) %>% 
  rowwise() %>% 
  mutate( nacq = f_supply(cbg, n0=1.0) )

df_nsupply %>% 
  ggplot(aes(x=cbg, y=nacq)) +
  geom_line(color="red", size=1) +
  ylim(0,1) +
  geom_hline( yintercept=1.0, linetype="dashed" ) +
  theme_classic() +
  labs(title = "N supply", x = expression(C[bg]), y = expression(N[acq]))
```

### N demand
N demand $N_\text{req}$ is governed by the amount of new biomass production, which is a function of the N:C stoichiometry ($r$) and the gross primary productivity (respiration is ignored here). The latter is itself a saturating function of leaf surface area ($L$). For mathematical tractability, let's use another time the Monod function to describe this:
$$
N_\text{req} = r P_0 \frac{L}{L+K_L}
$$
$P_0$ is the light use efficiency multiplied by incident photosynthetically active radiation. $K_L$ is the leaf half-saturation constant.
```{r echo=FALSE, message=FALSE, warning=FALSE}
prod <- function(leafarea, ppfd, lue){
  kl <- 50
  ppfd * lue * leafarea / (leafarea + kl )
}
f_ndemand <- function( ..., r_cton_plant ){
  (1/r_cton_plant) * prod(...)
}
r_cton_plant <- 30
df_ndemand <- tibble( leafarea = seq(0, 350, length.out = 100) ) %>% 
  rowwise() %>% 
  mutate( ndemand = f_ndemand(leafarea, ppfd=100, lue=1, r_cton_plant = r_cton_plant) )

df_ndemand %>% 
  ggplot(aes(x=leafarea, y=ndemand)) +
  geom_line(color="red", size=1) +
  # geom_hline( yintercept=1.0, linetype="dashed" ) +
  theme_classic() +
  labs(title = "N demand", x = expression(L), y = expression(N[req]))
```

By setting $N_\text{supply} = N_\text{demand}$, the root:shoot surface area ratio is given by the following expression:
$$
\frac{1}{\xi} = \frac{1}{K_R} \left[ L \left( \frac{N_0}{rP_0}-1 \right) + K_L \left( \frac{N_0}{rP_0} \right) \right]
$$
In steady state, the amount of N that becomes available from net mineralisation ($N_0$) must be equal to the amount of N added to the litter and soil through litterfall. Let's further assume that the available light is fully absorbed, then $N_0 = r P_0 + N_\text{in}$, where $N_\text{in}$ is the external N input. Plugging this into the equation above leads to:
$$
\frac{1}{\xi} = \frac{1}{K_R} \left[ L  \frac{N_\text{in}}{rP_0} + K_L \left( 1 + \frac{N_\text{in}}{rP_0} \right) \right]
$$
This is instructive because, we can immediately see, tha upon an increase in the external N inputs, the root:shoot ratio $\xi$ should go down. In contrast, if the the light use efficiency increases under elevated CO2 for example, thus $P_0$ increases, then the root:shoot ratio should go up.


## Solution

For a given total biomass ($C_\text{tot}$), both $N_\text{demand}$ and $N_\text{supply}$ can be expressed as a function of $\xi$ and set equal. To do so, $L$ and $C_\text{bg}$ are to be expressed as functions of $\xi$. 

$$
C_\text{ag} = L/\sigma
$$

$$
\xi = \frac{C_\text{bg}}{C_\text{ag}} = \frac{C-L/\sigma}{L/\sigma}
$$
Hence
$$
L = \frac{\sigma C}{\xi + 1}
$$
and
$$
C_\text{bg} = \frac{C}{1 + \frac{1}{\xi}}
$$

Or the same thing expressed in terms of the mass of foliage C as a fraction of total C:
$$
\alpha = \frac{C_\text{ag}}{C} = \frac{L/\sigma}{C}
$$
Hence
$$
L = \alpha \sigma C
$$
and
$$
C_\text{bg} = (1 - \alpha) C
$$

```{r echo=FALSE, message=FALSE, warning=FALSE}
# leaf area and belowground C as a function of root:shoot ratio
calc_leafarea_rrs <- function(rrootshoot, ctot, sla){
  (sla * ctot)/(rrootshoot + 1)
}
calc_cbg_rrs <- function(rrootshoot, ctot){
  ctot / (1 + 1/rrootshoot)
}

# leaf area and belowground C as a function of fraction in shoots (alpha)
calc_leafarea_alpha <- function(alpha, ctot, sla){
  alpha * sla * ctot
}
calc_cbg_alpha <- function(alpha, ctot){
  (1-alpha) * ctot
}

sla <- 0.1
ctot <- 700
ppfd <- 100
lue  <- 1
n0   <- 2

df_rrs <- tibble( rrootshoot = seq(0, 5, length.out = 100) ) %>% 
  rowwise() %>% 
  mutate( ndemand = f_ndemand( calc_leafarea_rrs( rrootshoot, ctot, sla ), ppfd=ppfd, lue=lue, r_cton_plant = r_cton_plant),
          nsupply = f_supply(  calc_cbg_rrs(      rrootshoot, ctot ), n0=n0) )

df_alpha <- tibble( alpha = seq(0, 1, length.out = 100) ) %>% 
  rowwise() %>% 
  mutate( ndemand = f_ndemand( calc_leafarea_alpha( alpha, ctot, sla ), ppfd=ppfd, lue=lue, r_cton_plant = r_cton_plant),
          nsupply = f_supply(  calc_cbg_alpha(      alpha, ctot ), n0=n0) )


## Numerical search for solution (because EGU is in four days)
setzero_rrs <- function(rrs, ctot, n0, ppfd, lue, r_cton_plant, sla){
  nsupply <- f_supply(  calc_cbg_rrs(      rrs, ctot ), n0=n0 )
  ndemand <- f_ndemand( calc_leafarea_rrs( rrs, ctot, sla ), ppfd=ppfd, lue=lue, r_cton_plant = r_cton_plant )
  out <- nsupply - ndemand
  return(out)
}
setzero_alpha <- function(alpha, ctot, n0, ppfd, lue, r_cton_plant, sla){
  nsupply <- f_supply(  calc_cbg_alpha(      alpha, ctot ), n0=n0 )
  ndemand <- f_ndemand( calc_leafarea_alpha( alpha, ctot, sla ), ppfd=ppfd, lue=lue, r_cton_plant=r_cton_plant )
  out <- nsupply - ndemand
  return(out)
}

root_rrs   <- uniroot( function(x) setzero_rrs( x, ctot, n0, ppfd, lue, r_cton_plant, sla ),   interval=c(0,5) )
root_alpha <- uniroot( function(x) setzero_alpha( x, ctot, n0, ppfd, lue, r_cton_plant, sla ), interval=c(0,1) )

df_rrs %>% 
  tidyr::gather(var, nflux, c(ndemand, nsupply)) %>% 
  ggplot(aes(x=rrootshoot, y=nflux, color=var)) +
  geom_line(size=1) +
  theme_classic() +
  labs(title = "N supply = N demand", x = expression(xi), y = "N flux", color = "" ) +
  geom_point( aes( x = root_rrs$root, y = f_supply( calc_cbg_rrs( root_rrs$root, ctot ), n0=n0) ), color="black", size=2 )

df_alpha %>% 
  tidyr::gather(var, nflux, c(ndemand, nsupply)) %>% 
  ggplot(aes(x=alpha, y=nflux, color=var)) +
  geom_line(size=1) +
  theme_classic() +
  labs(title = "N supply = N demand", x = expression(alpha), y = "N flux", color = "" ) +
  geom_point( aes( x = root_alpha$root, y = f_supply( calc_cbg_alpha( root_alpha$root, ctot ), n0=n0) ), color="black", size=2 )
```

With this balance, defined by $\alpha$, the N supply is equal to the N demand, imposed by productivity and the C:N ratio of new production. We can express the ratio of newly acquired C to N as a function of $\alpha$. The root (intersection) found above should be exactly where this function is equal to the predefined C:N ratio of biomass.
```{r}
alpha <- root_alpha$root
leafarea <- ctot * alpha * sla
nsupply <- f_supply(  calc_cbg_alpha(      alpha, ctot ), n0=n0 )
ndemand <- f_ndemand( calc_leafarea_alpha( alpha, ctot, sla ), ppfd=ppfd, lue=lue, r_cton_plant=r_cton_plant )
cfix    <- prod( leafarea, ppfd, lue )
print(cfix/nsupply)
```

## Dynamic CN-model

Above equations can also be implemented in a simple dynamical model that additionally accounts for soil C and N turnover. The results from a spinup to equilibrium look like this (suggesting that there is a physically meaningful solution). In this case (last plot at the very bottom), the ratio of $C_l/C$ is around 0.83.

```{r echo=FALSE}
# lue  <- 1
# ppfd <- 70 # for example with MM
# eff  <- 0.7
# resp <- 0.1
# r_cton_plant <- 30
# f_unavoidable <- 0.1
# k_noloss <- 100
# kmm_prod <- 100

ctot <- 700
ppfd <- 100
lue  <- 1
n_in <- 1.0

## other parameters defined above
r_cton_plant <- 30
r_cton_soil <- 10
tau_plant <- 10
tau_soil_c <- 50
tau_soil_n <- 150
tau_labl   <- 0.5
sla        <- 0.1
eff        <- 1.0

r_ntoc_plant <- 1/r_cton_plant
r_ntoc_soil  <- 1/r_cton_soil

## Simulation settings
ntsteps <- 2000

## initialise output variables
out_cplant_ag <- rep(NA, ntsteps)
out_nplant_ag <- rep(NA, ntsteps)
out_cplant_bg <- rep(NA, ntsteps)
out_nplant_bg <- rep(NA, ntsteps)
out_csoil     <- rep(NA, ntsteps)
out_nsoil     <- rep(NA, ntsteps)
out_clabl     <- rep(NA, ntsteps)
out_nlabl     <- rep(NA, ntsteps)

## plant, aboveground
cplant_ag <- 150
nplant_ag <- cplant_ag * r_ntoc_plant

## plant, belowground
cplant_bg <- 150
nplant_bg <- cplant_bg * r_ntoc_plant

## soil
csoil <- 1000
nsoil <- csoil * r_ntoc_plant * tau_soil_n/tau_soil_c

## plant labile pools
nlabl <- 0
clabl <- 0

## to avoid numerical oscillation
calc_turnover <- function( c0, tau, dt = 1.0 ){
  c0 * (1.0 - exp(-1/tau * dt ) )
}

for (it in 1:ntsteps){

  # ## step increase in light use efficiency ~ CO2 fertilisation
  # if (it==1000) lue <- 1.2 * lue

  ## Soil turnover
  csoil_turnover <- calc_turnover( csoil, tau_soil_c )
  csoil          <- csoil - csoil_turnover

  nsoil_turnover <- calc_turnover( nsoil, tau_soil_n )
  nsoil          <- nsoil - nsoil_turnover

  ## Net mineralisation
  netmin         <- nsoil_turnover + n_in

  ## Plant turnover, needs to be after acquisition and before new balance evaluation
  cturnover_ag <- calc_turnover( cplant_ag, tau_plant )
  cplant_ag    <- cplant_ag - cturnover_ag

  nturnover_ag <- calc_turnover( nplant_ag, tau_plant)
  nplant_ag    <- nplant_ag - nturnover_ag

  cturnover_bg <- calc_turnover( cplant_bg, tau_plant )
  cplant_bg    <- cplant_bg - cturnover_bg

  nturnover_bg <- calc_turnover( nplant_ag, tau_plant)
  nplant_ag    <- nplant_ag - nturnover_bg

  csoil <- csoil + cturnover_ag + cturnover_bg
  nsoil <- nsoil + nturnover_ag + nturnover_bg

  ## update total biomass with (allocatable) labile C and N 
  ctot <- cplant_ag + cplant_bg + clabl
  ntot <- nplant_ag + nplant_bg + nlabl
  print( paste( "C:N ratio of plant:", ctot/ntot ) )
  
  ## Get balanced allocation
  root_alpha <- uniroot( function(x) setzero_alpha( x, ctot, netmin, ppfd, lue, r_cton_plant, sla ), interval=c(0,1) )
  
  ## redesign the plant (immediate effect assumption)
  clabl <- 0
  nlabl <- 0
  cplant_ag <- root_alpha$root * ctot
  aleaf <- sla * cplant_ag
  cplant_bg <- (1 - root_alpha$root) * ctot
  
  nplant_ag <- cplant_ag * r_ntoc_plant
  nplant_bg <- cplant_bg * r_ntoc_plant
  
  # nsupply <- f_supply( calc_cbg_alpha( alpha, ctot ), n0=n0 )
  nlabl <- f_supply( cplant_bg, n0=netmin ) #+ nlabl
  clabl <- prod( aleaf, ppfd, lue ) #+clabl
  print( paste( "C:N ratio of labile:", clabl/nlabl ) )

  # ## short-cut: by-passing soil
  # nlabl <- f_noloss( cplant_bg ) * r_ntoc_plant * (cplant_bg + cplant_ag) / tau_plant  #+ nlabl

  # ## limit allocatable C depending on labile N, c_alloc is used also for growth respiration
  # c_alloc <- min( nlabl / (eff * r_ntoc_plant), clabl )
  # n_alloc <- eff * c_alloc * r_ntoc_plant

  # ## Get optimal fraction to leaves
  # fleaf <- findroot_eval_cnbalance( (c_alloc * eff + cplant_ag + cplant_bg), netmin, lue, ppfd )
  # # print(fleaf)

  # ##----------
  # ## Visualise imbalance curve
  # ctot <- (cplant_ag + cplant_bg + eff * c_alloc)
  # par(mfrow=c(1,1))
  # curve( eval_cnbalance( x, ctot, netmin, lue, ppfd ), from = 0, to = 1.0, col="red", ylim=c(-20,100) )
  # abline( h=0, lty=3 )
  # abline( v = fleaf, lty = 2 )

  # ## production curve
  # curve( prod( f_leaf_of_c_ag_tot( x ) * x, ppfd, lue ), from = 0, to = ctot ) # x is total aboveground
  # points( fleaf * ctot, prod( f_leaf_of_c_ag_tot( fleaf * ctot ) * fleaf * ctot, ppfd, lue ), pch=16, col="red" )

  # ## N uptake curve
  # curve( netmin * f_noloss( x ), from = 0, to = ctot ) # x is total belowground
  # points( (1.0 - fleaf) * ctot, netmin * f_noloss( (1.0 - fleaf) * ctot ), pch=16, col="red" )
  # ##----------

  ## Additional check only necessary for legacy allocation setup (Version A)
  # if ( eval_cnbalance( 0, ctot, netmin, lue, ppfd ) > 0.0 ){
  #   ## all to roots
  #   fleaf <- 0.0
  # } else if ( eval_cnbalance( 1, ctot, netmin, lue, ppfd ) < 0.0 ){
  #   ## all to leaves
  #   fleaf <- 1.0
  # } else {
  #   ## find allocation fraction to leaves so that return next year matches (eff * r_cton_plant)
  #   out_root <- uniroot( function(x) eval_cnbalance( x, (cplant_ag + cplant_bg + eff * c_alloc), netmin, lue, ppfd ), interval=c(1e-12,1.0) )
  #   fleaf <- out_root$root
  # }

  # ## update pools
  # clabl <- clabl - c_alloc
  # nlabl <- nlabl - n_alloc

  # ## labile pool decay
  # clabl <- clabl - calc_turnover( clabl, tau_labl)
  # nlabl <- nlabl - calc_turnover( nlabl, tau_labl)

  ##---------------------------------------------------------------------
  # ## VERSION A: fleaf only acts on new growth
  # ## allocate C, N follows from r_ntoc_plant
  # cturnover_ag <- cplant_ag / tau_plant
  # c_alloc_ag   <- fleaf * c_alloc * eff
  # cplant_ag    <- cplant_ag + c_alloc_ag - cturnover_ag

  # nturnover_ag <- nplant_ag / tau_plant
  # n_alloc_ag   <- c_alloc_ag * r_ntoc_plant
  # nplant_ag    <- nplant_ag + n_alloc_ag - nturnover_ag

  # cturnover_bg <- cplant_bg / tau_plant
  # c_alloc_bg   <- (1.0 - fleaf) * c_alloc * eff
  # cplant_bg    <- cplant_bg + c_alloc_bg - cturnover_bg

  # nturnover_bg <- nplant_bg / tau_plant
  # n_alloc_bg   <- c_alloc_bg * r_ntoc_plant
  # nplant_bg    <- nplant_bg + n_alloc_bg - nturnover_bg
  ##---------------------------------------------------------------------

  # ##---------------------------------------------------------------------
  # ## VERSION B: fleaf redistributes entire plant (to avoid legacy in mal-allocation)
  # ## allocate C, N follows from r_ntoc_plant
  # cplant_ag    <- fleaf * (c_alloc * eff + cplant_ag + cplant_bg)
  # cplant_bg    <- (1.0 - fleaf) * (c_alloc * eff + cplant_ag + cplant_bg)
  # ##---------------------------------------------------------------------

  ## gater output variables
  out_cplant_ag[it] <- cplant_ag
  out_nplant_ag[it] <- nplant_ag
  out_cplant_bg[it] <- cplant_bg
  out_nplant_bg[it] <- nplant_bg
  out_csoil[it]     <- csoil
  out_nsoil[it]     <- nsoil
  out_clabl[it]     <- clabl
  out_nlabl[it]     <- nlabl

}
##---------------------------------------------------------
plot( 1:ntsteps, out_cplant_ag, type = "l" )
plot( 1:ntsteps, out_cplant_bg, type = "l" )
plot( 1:ntsteps, out_csoil, type = "l" )
plot( 1:ntsteps, out_nsoil, type = "l" )
plot( 1:ntsteps, out_clabl, type = "l" )
plot( 1:ntsteps, out_nlabl, type = "l" )
plot( 1:ntsteps, out_cplant_ag/(out_cplant_bg+out_cplant_ag), type = "l" )

print("Steady state f_ag (fraction of aboveground to total plant C):")
print( cplant_ag[ntsteps]/(cplant_bg[ntsteps]+cplant_ag[ntsteps]) )

print("Steady state total plant C:")
print( cplant_bg[ntsteps]+cplant_ag[ntsteps] )

print("Steady state net mineralisation:")
print(netmin)

print("Expected steady state net mineralisation:")
print( (cplant_ag + cplant_bg) / (r_cton_plant * tau_plant) + n_in )
```


<!-- It tends to achieve a balance between supply of N through acquisition pathways and the demand imposed by new growth and its respective C:N ratio. -->
<!-- $$ -->
<!-- \nacq (C_r) = y/\rcton [ P(C_l) - r(C_l+C_r) ] \; \; \; \; \; \; (1) -->
<!-- $$ -->
<!-- $P$ is the C assimilation rate and is a function of leaf mass ($C_l$), $y$ is the growth efficiency, $r$ a respiration coefficient, implying that plant respiration scales with its size, $C_l$ and $C_r$ are the leaf carbon and root pool sizes respectively, and $N_{\text{acq}}$ is the the N acquisition flux and is a function of the root mass ($C_r$).  -->

<!-- - The size of the above-ground pool $C_l$ (each pool consists of C and N) determines the assimilation rate of C ($P$) and has declining marginal returns towards increasing $C_l$. -->
<!-- $$ -->
<!-- P(C_l) = I \varepsilon \left(1 - e^{-k_b \sigma C_{\text{l}}} \right) -->
<!-- $$ -->
<!-- Here, $I$ is PPFD, $\varepsilon$ is the light use efficiency, $k_b$ is the light extinction parameter, and $\sigma$ is the specific leaf area. -->
<!-- To simplify the mathematics, I've tried to model this with Michaelis-Menten as an alternative: -->
<!-- $$ -->
<!-- P(C_l) = I \varepsilon \frac{C_l}{C_l + K_P} -->
<!-- $$ -->

<!-- ```{r echo=FALSE} -->
<!-- prod <- function( cleaf, ppfd, lue ){ -->
<!--   # kbeer         = 0.5 -->
<!--   # sla           = 0.1 -->
<!--   #prod <- ppfd * lue * ( 1.0 - exp( - kbeer * sla * cleaf ) ) -->

<!--   kmm_prod <- 100 -->
<!--   prod <- ppfd * lue * cleaf / ( cleaf + kmm_prod ) -->
<!--   return(prod) -->
<!-- } -->
<!-- ## Example -->
<!-- curve( prod( x, ppfd = 100, lue = 1 ), from = 0, to = 350, xlab = expression(italic(C[l])), ylab = expression(italic(P)) ) -->
<!-- ``` -->


<!-- - The size of the below-ground pool ($C_r$) determines the fraction of net N mineralisation $\netmin$ that is acquired by the plant ($\nacq$). As for $P$, the function $f(C_r)$ represents declining marginal returns in $\nacq$ with increasing $C_r$: -->
<!-- $$ -->
<!-- \nacq = f(C_r) \netmin   \\ -->
<!-- f(C_r) = (1 - u)  (1 - e^{-k_r  C_r } )  -->
<!-- $$ -->
<!-- Here, $u$ is the unaccessible fraction of net mineralisation, i.e. N losses that are not subject to plant control. -->
<!-- Again, in order to simplify mathematics, I've tried to model this with Michaelis-Menten as an alternative: -->
<!-- $$ -->
<!-- f(C_r) = (1-u) \frac{C_r}{C_r + K_N} -->
<!-- $$ -->

<!-- ```{r echo=FALSE} -->
<!-- f_noloss <- function( croot ){ -->
<!--   f_unavoidable = 0.1 -->
<!--   k_noloss      = 100 -->
<!--   out <- (1.0 - f_unavoidable) * croot / (croot + k_noloss ) -->
<!--   # f_avl <- min( 1.0, croot / 30 ) -->
<!--   # out   <- (1.0 - f_unavoidable) * (1.0 - exp( - 0.01 * croot ) ) -->
<!--   return(out) -->
<!-- } -->
<!-- ## Example -->
<!-- netmin <- 1.0 -->
<!-- curve( netmin * f_noloss(x), from = 0, to = 350, ylim=c(0,1), xlab = expression(italic(C[r])), ylab = expression(italic(N[acq])) ) -->
<!-- abline( h=(1-0.1), lty=3 ) -->
<!-- abline( h=netmin, lty=2 ) -->
<!-- ``` -->

<!-- ## Example -->

<!-- For given $I$, $\varepsilon$, $N_{\text{min}}$, a balance of above-to-belowground pool size $a = C_l:C_r$ can be found that satisfies Eq. (1). The imbalance can be expressed as -->
<!-- $$ -->
<!-- f((1-a)C)\netmin = y/\rcton\; [ P(aC) - rC ] -->
<!-- $$ -->
<!-- Here, $C=C_l+C_r$ and $a=C_l/C$. Below is an example plot for the imbalance term as a function of $a$.  -->

<!-- ```{r echo=FALSE} -->
<!-- eval_cnbalance <- function( f_ag, ctot, netmin, lue, ppfd ){ -->
<!--   eff  = 0.7 -->
<!--   resp = 0.1 -->
<!--   r_cton_plant = 30 -->

<!--   ## version where not all aboveground C is leaves -->
<!--   # out <- (( eff  * ( prod( f_leaf_of_c_ag_tot( f_ag * ctot ) * f_ag * ctot, ppfd, lue ) - resp * ctot ) ) / -->
<!--   #        ( netmin * f_noloss( (1.0 - f_ag) * ctot ) ) ) - r_cton_plant -->

<!--   ## version where all aboveground C is leaves -->
<!--   # out <- (( eff  * ( prod( f_ag * ctot, ppfd, lue ) - resp * ctot ) ) / -->
<!--   #        ( netmin * f_noloss( (1.0 - f_ag) * ctot ) ) ) - r_cton_plant -->

<!--   ## this works much better than the ratio calculated above -->
<!--   supply <- f_noloss((1.0 - f_ag) * ctot ) * netmin -->
<!--   demand <- eff * ( prod( f_ag * ctot, ppfd, lue ) - resp * ctot ) / r_cton_plant -->
<!--   out <- supply - demand -->

<!--   return( out ) -->
<!-- } -->

<!-- findroot_eval_cnbalance <- function( ... ){ -->

<!--   if ( eval_cnbalance( 0.0, ... ) < 0.0 ){ -->
<!--     ## all to roots -->
<!--     fleaf <- 0.0 -->
<!--   } else if ( eval_cnbalance( 1.0, ... ) > 0.0 ){ -->
<!--     ## all to leaves -->
<!--     fleaf <- 1.0 -->
<!--   } else { -->
<!--     ## find allocation fraction to leaves so that return next year matches (eff * r_cton_plant) -->
<!--     out_root <- uniroot( function(x) eval_cnbalance( x, ... ), interval=c(0,1.0) ) -->
<!--     fleaf <- out_root$root -->
<!--   } -->
<!--   return(fleaf) -->
<!-- } -->

<!-- lue  = 1 -->
<!-- # ppfd = 60 # for example with exponentials -->
<!-- ppfd = 100 # for example with MM -->
<!-- netmin = 0.8 -->
<!-- ctot = 350 -->

<!-- ## Imbalance as a function of the C_l:C -->
<!-- curve( eval_cnbalance( x, ctot, netmin, lue, ppfd ), from = 0, to = 1.0, col="red", xlab=expression(italic(C[l]/C)), ylab=expression(Imbalance) ) #, ylim=c(-20,100) -->
<!-- abline( h=0, lty=3 ) -->
<!-- out_root <- findroot_eval_cnbalance( ctot, netmin, lue, ppfd )  # uniroot( function(x) eval_cnbalance( x, ctot, netmin, lue, ppfd ), interval=c(0,1.0) ) -->
<!-- abline( v = out_root, lty = 2 ) -->
<!-- ``` -->

<!-- ## Steady-state solution -->

<!-- In steady state, the total stock of N in the soil is constant. This implies that net mineralisation equals total N inputs through litterfall and N deposition (assuming deposited N bypasses plants). For the sake of simplicity, N fixation and resorption are ignored here. -->
<!-- $$ -->
<!-- \netmin = \frac{C}{\rcton\tau} + \nin -->
<!-- $$ -->
<!-- Here, $\tau$ is the residence time of C and N in the plant pool; $\rcton$ is the C:N ratio of plant biomass; and $\nin$ is the external input of N, subsuming atmospheric deposition and weathering. Using this, the steady-state solution of the coupled soil-plant system can be found, without relying on prescribed net mineralisation rates, with the balanced-growth condition from above: -->
<!-- $$ -->
<!--  f((1-a)C)(\frac{C}{\rcton\tau} +\nin) =  y / \rcton \; [ P(aC) - rC ]\; \; \; \;\; \; \; \;(2) -->
<!-- $$ -->
<!-- and the plant C blance (C assimilation equals respiration plus new growth, replacing litterfall in steady-state): -->
<!-- $$ -->
<!-- P(aC) - (\tau/y + r)\;C = 0 \; \; \; \; \; \; \; \;(3) -->
<!-- $$ -->
<!-- This set of two equations should in theory allow us to cancel $C$ and solve for $a$ as a function of environmental conditions: $a = f(\varepsilon, I, \nin)$. Note that $\varepsilon$ is not treated here as a parameter because it's itself a function of environmental conditions. Let's not try to get the analytical solution here. It will involve Weibull functions anyways, and these have to be solved numerically. Let's just throw it at `nleqslv()`. -->

<!-- ```{r} -->
<!-- library(nleqslv) -->

<!-- ## Equation system as a function returning a vector of two values, both of which are to be set to zero by the optimiser -->
<!-- fn <- function( par ){ -->
<!--   eff  <- 0.7 -->
<!--   resp <- 0.1 -->
<!--   r_cton_plant <- 30 -->
<!--   tau_plant <- 10 -->
<!--   n_in <- 1.0 -->

<!--   out <- numeric(2) -->
<!--   ctot <- par[1] -->
<!--   f_ag <- par[2] -->

<!--   out[1] <- eval_cnbalance( f_ag, ctot, ( ctot/(r_cton_plant * tau_plant) + n_in), lue, ppfd ) -->
<!--   out[2] <- prod( f_ag * ctot, ppfd, lue ) - ctot * ( tau_plant/eff + resp ) -->

<!--   return(out) -->
<!-- } -->

<!-- steadystate <- nleqslv( c(500, 0.5), fn, control = list( allowSingular=TRUE ) ) -->

<!-- ## test -->
<!-- ctot <- steadystate$x[[1]] -->
<!-- f_ag <- steadystate$x[[2]] -->
<!--   eff  <- 0.7 -->
<!--   resp <- 0.1 -->
<!--   r_cton_plant <- 30 -->
<!--   tau_plant <- 10 -->
<!--   n_in <- 1.0 -->
<!--   print(paste("should be zero:", eval_cnbalance( f_ag, ctot, ( ctot/(r_cton_plant * tau_plant) + n_in), lue, ppfd ) )) -->
<!--   print(paste("should be zero:", prod( f_ag * ctot, ppfd, lue ) - ctot * ( tau_plant/eff + resp ) )) -->
<!-- ``` -->

<!-- ### My problem -->

<!-- Something is apparently not right here. I've tried to solve this system of two non-linear equations numerically but the solution I get is nonsense. A problem might be that I cannot constrain the solutions to a (phyically meaningful) interval, $0<a<1$ in this case. But the nature of the problem shouldn't require this anyways. So I am asking myself if the second equation above (Eq. 3) is appropriate or if something else is wrong. -->

<!-- So far, I haven't attempted to analytically solve this. Feel free if you like... -->

<!-- ```{r eval=FALSE, echo=FALSE} -->

<!-- # ## enther this in wolfram alpha: -->
<!-- # ((1 - u) * (1 - y) * x / ((1 - y) * x + l)) * ( x/(s*t) + d) - e * ( (I * p * x * y / (x * y + k)) - r * x ) / s = 0 -->
<!-- # (I * p * x * y / (x * y + k)) - (t/e + r) * x = 0 -->

<!-- lue  <<- 1 -->
<!-- ppfd <<- 100 # for example with MM -->
<!-- eff  <<- 0.7 -->
<!-- resp <<- 0.1 -->
<!-- r_cton_plant <<- 30 -->
<!-- f_unavoidable <<- 0.1 -->
<!-- k_noloss <<- 100 -->
<!-- kmm_prod <<- 100 -->
<!-- n_in <<- 1.0 -->
<!-- tau_plant <<- 10 -->

<!-- fn <- function( x ){ -->

<!--   ctot <- x[1] -->
<!--   f_ag <- x[2] -->
<!--   out <- numeric(2) -->

<!--   supply <- f_noloss((1.0 - f_ag) * ctot ) * netmin -->
<!--   demand <- eff * ( prod( f_ag * ctot, ppfd, lue ) - resp * ctot ) / r_cton_plant -->
<!--   out[1] <- supply - demand -->

<!--   out[2] <- prod( f_ag * ctot, ppfd, lue ) - (tau_plant / eff + resp) * ctot -->

<!--   return(out) -->
<!-- } -->

<!-- ## Does not work, maybe I'm looking in an implausible range of values -->
<!-- library(pracma) -->
<!-- solution_pracma <- pracma::fsolve( fn, c(230.8476, 0.8366655) ) -->

<!-- library(nleqslv) -->
<!-- solution_nleqslv <- nleqslv::nleqslv( c(230.8476, 0.8366655), fn ) -->

<!-- print( "Steady-state solution:" ) -->
<!-- print( paste0( "C_l: ", solution_pracma$x[1] * solution_pracma$x[2] )) -->
<!-- print( paste0( "C_r: ", solution_pracma$x[1] * (1-solution_pracma$x[2]) )) -->
<!-- ``` -->

<!-- <!-- ```{r} --> -->
<!-- <!-- c_wood <- function( ctot ){ --> -->
<!-- <!--   par = 50 --> -->
<!-- <!--   sqrt( ctot^2 + par ) - sqrt(par) --> -->
<!-- <!-- } --> -->

<!-- <!-- c_leaf <- function( ctot ){ --> -->
<!-- <!--   kl = 0.3 --> -->
<!-- <!--   c_leaf_max = 20 --> -->
<!-- <!--   c_leaf_max * (1 - exp(-kl*ctot)) --> -->
<!-- <!-- } --> -->

<!-- <!-- f_leaf_of_c_ag_tot <- function( c_ag_tot ){ --> -->
<!-- <!--   ifelse( c_ag_tot==0, 1.0, c_leaf( c_ag_tot ) / (c_leaf( c_ag_tot ) + c_wood( c_ag_tot )) ) --> -->
<!-- <!-- } --> -->
<!-- <!-- ``` --> -->

<!-- <!-- ## Static CN-model for one state------------------------------- --> -->
<!-- <!-- lue  = 1 --> -->
<!-- <!-- ppfd = 100 --> -->
<!-- <!-- navl = 0.8 --> -->
<!-- <!-- ctot = 350 --> -->

<!-- <!-- curve( c_wood(x), from = 0, to = 50 ) --> -->
<!-- <!-- curve( c_leaf(x), from = 0, to = 50, add = TRUE ) --> -->
<!-- <!-- curve( f_leaf_of_c_ag_tot(x), from = 0, to = ctot ) --> -->

<!-- <!-- curve( prod( f_leaf_of_c_ag_tot( x ) * x, ppfd, lue ), from = 0, to = ctot ) --> -->
<!-- <!-- curve( f_noloss, from = 0, to = ctot*2 ) --> -->

<!-- <!-- ## Visualise imbalance curve --> -->
<!-- <!-- par(mfrow=c(1,1)) --> -->
<!-- <!-- curve( eval_cnbalance( x, ctot, navl, lue, ppfd ), from = 0, to = 1.0, col="red", ylim=c(-20,100) ) --> -->
<!-- <!-- abline( h=0, lty=3 ) --> -->
<!-- <!-- out_root <- uniroot( function(x) eval_cnbalance( x, ctot, navl, lue, ppfd ), interval=c(0,1.0) ) --> -->
<!-- <!-- abline( v = out_root$root, lty = 2 ) --> -->

<!-- <!-- ## production curve --> -->
<!-- <!-- curve( prod( f_leaf_of_c_ag_tot( x ) * x, ppfd, lue ), from = 0, to = ctot ) # x is total aboveground --> -->
<!-- <!-- points( out_root$root * ctot, prod( f_leaf_of_c_ag_tot( out_root$root * ctot ) * out_root$root * ctot, ppfd, lue ), pch=16, col="red" ) --> -->

<!-- <!-- ## N uptake curve --> -->
<!-- <!-- curve( navl * f_noloss( x ), from = 0, to = ctot ) # x is total belowground --> -->
<!-- <!-- points( (1.0 - out_root$root) * ctot, navl * f_noloss( (1.0 - out_root$root) * ctot ), pch=16, col="red" ) --> -->



